---
title: "Causal Inference Project: G-Computation for Adversarial Multi-Agent Trust Games"
author: "Harshit Soni (hs5666), Gabriel Nixon Raj (gr2513)"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    fig_caption: true
    keep_tex: true
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6,
  fig.align = "center",
  cache = FALSE
)

library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(purrr)
library(knitr)
library(kableExtra)

set.seed(42)
```

```{r config, include=FALSE}
# ============================================================
# CONFIGURATION — Toggle DGP on/off
# ============================================================
# Set RUN_DGP = TRUE to regenerate simulation data from scratch.
# Set RUN_DGP = FALSE to load existing DGP output and run analysis only.

RUN_DGP     <- FALSE
N_SEEDS     <- 34
SEED_START  <- 1
N_BOOT      <- 1000
ALPHA       <- 0.05
PROJECT_DIR <- "."  # Set to your MRTT project root if different
PYTHON_CMD  <- "python"
TIME_BUDGET <- 8  # hours (only used if RUN_DGP = TRUE)
```


# Motivation

Adversarial manipulation of human decision-making is a growing concern in domains where
automated agents interact repeatedly with human users. Dezfouli, Nock, and Dayan (2020)
demonstrated that reinforcement learning agents can discover and exploit systematic
biases in human sequential decision-making within a multi-round trust game (MRTT).
Their work raises a natural causal question: *does the structure of information sharing
among adversarial agents causally affect their ability to exploit a human decision-maker?*

We extend the Dezfouli et al. framework by introducing a *world* variable $W \in \{0, 1\}$
that controls whether adversarial agents can observe each other's actions and states.
In $W = 0$ (independent), each agent interacts with the investor through a shared
learner but cannot observe other agents' strategies. In $W = 1$ (communication),
agents observe each other, potentially enabling coordinated exploitation.
Crucially, even under $W = 0$, agents are not truly independent: they are coupled
through the shared investor who learns from all agents simultaneously, creating
indirect interference through a common cause.

We chose g-computation (the parametric g-formula) as our primary estimator for
three reasons. First, our fully-specified structural causal model (SCM) directly
motivates g-computation as the natural identification strategy under sequential
ignorability. Second, g-computation has not been covered in course homeworks,
which focused on propensity score methods, instrumental variables, and BART.
Third, the g-formula allows us to cleanly demonstrate what happens when its key
assumption --- sequential ignorability --- holds versus when it is violated due
to unobserved latent states, a distinction built directly into our simulation
design.


# Assumptions

## Structural Assumptions

The SCM induces the following causal dependencies at each time step $t$:
$$h_t \rightarrow a_t \rightarrow x_t \rightarrow r_t \rightarrow h_{t+1}$$
with simultaneous wealth dynamics $a_t, x_t \rightarrow w_{t+1}$ and world-level
structural modifications $W \rightarrow h_{t,i}$ and $W \rightarrow x_{t,i}$.
The world variable $W$ modifies the structural laws governing agent behavior
rather than entering as an ordinary covariate.

## Sequential Ignorability

G-computation requires that at each time step $t$, the treatment assignment
is ignorable given the observed history:
$$Y_t(\bar{a}_t) \perp A_t \mid X_t$$
where $X_t$ includes all relevant state variables up to time $t$. In our
simulation, this holds *by construction* when the analyst observes the full
state including the RNN hidden states $h_t$, since these are the sufficient
statistics for the behavioral model's policy.

## Positivity

For each level of the covariates $X_t$, both treatment conditions must have
positive probability: $0 < P(W = 1 \mid X_t) < 1$. Since $W$ is randomized
by the experimenter (us) at the episode level and is independent of all
pre-treatment covariates, positivity holds trivially.

## SUTVA Considerations

The stable unit treatment value assumption is potentially violated because
agents within a game are coupled through the shared investor. Under $W = 0$,
interference operates indirectly through the investor's learning dynamics.
Under $W = 1$, direct agent-to-agent observation creates additional
interference channels. Our analysis operates at the episode level where $W$
is constant, sidestepping within-episode interference by treating the
entire episode trajectory as the outcome.

## Consistency

The consistency assumption requires that the observed outcome under treatment
$W = w$ equals the potential outcome $Y(w)$. Since $W$ is a structural
parameter of the simulation --- not a choice variable influenced by agent
behavior --- consistency holds by design.


# Design and Estimator

## G-Computation (Parametric G-Formula)

G-computation, introduced by Robins (1986), is a standardization-based
approach to causal inference under time-varying confounding. The key idea
is to model the outcome as a function of treatment and covariates, then
*standardize* by computing predicted outcomes under each treatment condition
for all units.

The procedure is:

1. **Fit an outcome model**: Estimate $\hat{E}[Y \mid W, X]$ using
   linear regression with treatment--covariate interactions.
2. **Predict potential outcomes**: For each unit $i$, compute
   $\hat{Y}_i(1) = \hat{E}[Y \mid W = 1, X_i]$ and
   $\hat{Y}_i(0) = \hat{E}[Y \mid W = 0, X_i]$.
3. **Estimate the ATE**: $\hat{\tau} = \frac{1}{n}\sum_{i=1}^{n}[\hat{Y}_i(1) - \hat{Y}_i(0)]$.

## Comparison with Related Methods

Unlike inverse-probability weighting (IPW), g-computation does not require
modeling the treatment assignment mechanism --- useful here since $W$ is
randomized and the "propensity score" is constant at 0.5. Unlike doubly-robust
estimators (AIPW), it relies entirely on correct specification of the outcome
model. This makes g-computation a clean test case: when the outcome model is
correctly specified (Setting A), the estimator is consistent; when it is
misspecified due to omitted confounders (Setting B), we can directly observe
the resulting bias.

## Implementation

We implement g-computation with linear outcome models including main effects
and treatment--covariate interactions. The covariate sets differ across settings:

- **Setting A (Full Information)**: Episode number, run-level RNN hidden state
  summaries (10 dimensions), and eval-phase behavioral statistics. This
  represents the scenario where the analyst has access to the agents' latent
  internal states.
- **Setting B (Behavioral Only)**: Episode number alone. This represents the
  realistic scenario (analogous to Dezfouli et al.'s human experiments) where
  the analyst observes only behavioral data and the agents' internal
  representations are latent.
- **Naive Baseline**: Simple difference in means with no covariate adjustment.


# Target Estimand

The average treatment effect (ATE) of inter-agent communication on
game outcomes:
$$\tau_{\text{ATE}} = E[Y(W = 1)] - E[Y(W = 0)]$$

We report the ATE for three outcome variables:

1. **Investor cumulative return**: Total wealth accumulated by the investor
   (learner) over an episode. This captures the net effect of adversarial
   communication on the investor's welfare.
2. **Fair agent mean repayment fraction**: Average repayment rate by the fair
   agent. A causal effect here indicates that communication changes how
   cooperative agents behave.
3. **Max agent mean repayment fraction**: Average repayment rate by the
   max (exploitative) agent. A causal effect here directly measures whether
   communication enables more effective exploitation strategies.


# Simulation Setup

## Data Generating Process

```{r dgp-run, eval=RUN_DGP}
# This chunk runs the Python MRTT simulation via the R driver.
# Only executes if RUN_DGP = TRUE in the config chunk above.

cat("Running DGP...\n")
cat(sprintf("Seeds: %d-%d, Time budget: %.1f hours\n",
            SEED_START, SEED_START + N_SEEDS - 1, TIME_BUDGET))

dgp_cmd <- sprintf(
  "Rscript run_dgp.R --n-seeds %d --seed-start %d --time-budget-hours %.1f --python %s",
  N_SEEDS, SEED_START, TIME_BUDGET, PYTHON_CMD
)
system(dgp_cmd, intern = FALSE)
```

```{r load-data}
# ============================================================
# Load and stitch DGP output
# ============================================================

dgp_files <- list.files(PROJECT_DIR,
                         pattern = "^mrtt_dgp_output.*\\.csv$",
                         full.names = TRUE)
stopifnot("No mrtt_dgp_output*.csv files found" = length(dgp_files) > 0)

dgp <- map_dfr(dgp_files, ~ read_csv(.x, show_col_types = FALSE)) %>%
  distinct(seed, world, episode, .keep_all = TRUE) %>%
  arrange(seed, world, episode)

cat(sprintf("Loaded %d rows: %d seeds × 2 worlds × %d episodes\n",
            nrow(dgp), n_distinct(dgp$seed),
            dgp %>% count(seed, world) %>% pull(n) %>% median()))
```

## Features Under Investigation

We explore two key features of g-computation:

1. **Reliance on sequential ignorability.** G-computation requires that all
   confounders between treatment and outcome are observed and conditioned upon.
   We test this by varying the analyst's information set: Setting A observes
   the full state (ignorability holds); Setting B omits the RNN hidden states
   (ignorability may be violated).

2. **Sensitivity to outcome model specification.** With randomized treatment,
   covariate adjustment should improve precision without introducing bias.
   We compare the g-computation estimates against a naive difference-in-means
   baseline to assess whether the added modeling complexity buys precision.

## DGP Description

The simulation uses the Multi-Round Trust Task (MRTT) architecture from
Dezfouli, Nock, and Dayan (2020), extended with:

- **Two adversarial agents** (one fair, one max) per game
- **A learning investor** (GRU-based RNN frozen from behavioral data, with a
  DQN Q-learner for investment allocation)
- **10,000 rounds per episode**, 100 training episodes per run
- **34 paired seeds**: each seed $s$ generates one W0 run and one W1 run with
  identical random initialization, providing a matched-pairs design

```{r dgp-summary}
dgp_summary <- dgp %>%
  group_by(world) %>%
  summarise(
    n_episodes = n(),
    n_seeds = n_distinct(seed),
    mean_inv_return = mean(investor_return),
    sd_inv_return = sd(investor_return),
    mean_fair_repay = mean(fair_1_repay),
    sd_fair_repay = sd(fair_1_repay),
    mean_max_repay = mean(max_1_repay),
    sd_max_repay = sd(max_1_repay),
    .groups = "drop"
  ) %>%
  mutate(world_label = ifelse(world == 1,
                               "W1 (Communication)", "W0 (Independent)"))

kable(dgp_summary %>%
        select(world_label, n_episodes, n_seeds,
               mean_inv_return, sd_inv_return,
               mean_fair_repay, sd_fair_repay,
               mean_max_repay, sd_max_repay),
      digits = 2,
      col.names = c("World", "Episodes", "Seeds",
                     "Mean Inv. Return", "SD Inv. Return",
                     "Mean Fair Repay", "SD Fair Repay",
                     "Mean Max Repay", "SD Max Repay"),
      caption = "Summary statistics by world condition") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE)
```

```{r fig-raw-distributions, fig.cap="Outcome distributions by world condition", fig.height=10}
raw_long <- dgp %>%
  select(seed, world, episode, investor_return, fair_1_repay, max_1_repay) %>%
  pivot_longer(c(investor_return, fair_1_repay, max_1_repay),
               names_to = "outcome", values_to = "value") %>%
  mutate(
    world_label = ifelse(world == 1, "W1 (Communication)", "W0 (Independent)"),
    outcome_label = case_when(
      outcome == "investor_return" ~ "Investor Cumulative Return",
      outcome == "fair_1_repay" ~ "Fair Agent Mean Repayment",
      outcome == "max_1_repay" ~ "Max Agent Mean Repayment"
    )
  )

ggplot(raw_long, aes(x = value, fill = world_label)) +
  geom_histogram(alpha = 0.5, bins = 40, position = "identity") +
  facet_wrap(~ outcome_label, scales = "free", ncol = 1) +
  scale_fill_manual(values = c("W0 (Independent)" = "#4575b4",
                                "W1 (Communication)" = "#d73027")) +
  labs(x = "Value", y = "Count", fill = "World") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom", strip.text = element_text(face = "bold"))
```

## Two Simulation Settings

```{r hidden-state-extraction}
# ============================================================
# Extract hidden-state features from per-run game_log.csv files
# (Setting A covariates)
# ============================================================

outputs_dir <- file.path(PROJECT_DIR, "outputs")
run_ids <- unique(dgp$run_id)

parse_hidden <- function(hs_str) {
  if (is.na(hs_str) || hs_str == "") return(rep(NA_real_, 5))
  as.numeric(strsplit(gsub("\\[|\\]", "", hs_str), ",\\s*")[[1]])
}

extract_run_features <- function(rid) {
  gl_path <- file.path(outputs_dir, rid, "game_log.csv")
  if (!file.exists(gl_path)) return(NULL)

  tryCatch({
    gl <- read_csv(gl_path, show_col_types = FALSE)

    # Per-agent behavioral summaries from eval simulation
    agent_stats <- gl %>%
      group_by(agent_name) %>%
      summarise(
        mean_repay = mean(repayment_pct, na.rm = TRUE),
        sd_repay = sd(repayment_pct, na.rm = TRUE),
        mean_invest = mean(investment, na.rm = TRUE),
        final_hs = last(rnn_hidden_state),
        .groups = "drop"
      )

    fair_hs <- parse_hidden(
      agent_stats %>% filter(agent_name == "fair_1") %>% pull(final_hs))
    max_hs <- parse_hidden(
      agent_stats %>% filter(agent_name == "max_1") %>% pull(final_hs))

    out <- tibble(
      run_id = rid,
      eval_repay_fair = agent_stats %>%
        filter(agent_name == "fair_1") %>% pull(mean_repay),
      eval_repay_max = agent_stats %>%
        filter(agent_name == "max_1") %>% pull(mean_repay),
      eval_sd_fair = agent_stats %>%
        filter(agent_name == "fair_1") %>% pull(sd_repay),
      eval_sd_max = agent_stats %>%
        filter(agent_name == "max_1") %>% pull(sd_repay),
      eval_invest = mean(gl$investment, na.rm = TRUE)
    )
    for (d in seq_along(fair_hs)) {
      out[[paste0("fair_h", d)]] <- fair_hs[d]
      out[[paste0("max_h", d)]] <- max_hs[d]
    }
    out
  }, error = function(e) NULL)
}

run_features <- map_dfr(run_ids, extract_run_features)
has_hidden <- nrow(run_features) > 0

if (has_hidden) {
  dgp_full <- dgp %>% left_join(run_features, by = "run_id")
  cat(sprintf("Hidden-state features extracted for %d / %d runs\n",
              nrow(run_features), length(run_ids)))
} else {
  dgp_full <- dgp
  cat("No per-run game_log.csv files found.\n")
  cat("Setting A will use episode-level covariates only.\n")
  cat("For full Setting A, ensure outputs/<run_id>/game_log.csv are accessible.\n")
}
```

**Setting A (Assumption Holds):** The analyst observes the full state, including
the RNN hidden state dimensions extracted from each run's evaluation simulation.
These hidden states capture the internal learned representation of each agent
--- the latent factors that mediate between treatment ($W$) and outcome. With
these observed, sequential ignorability is satisfied by construction: there are
no unobserved common causes of treatment and outcome that aren't conditioned on.

**Setting B (Assumption Violated):** The analyst observes only the behavioral
outcomes --- episode number and the outcome variables themselves. The RNN hidden
states are treated as unobserved. This mirrors the realistic scenario in
Dezfouli et al.'s human experiments, where the human's internal decision-making
state is latent. Sequential ignorability may fail because the hidden state
influences both how the investor allocates (related to treatment response) and
the episode outcomes.


# Simulation Results

```{r ground-truth}
# ============================================================
# Ground truth ATEs
# ============================================================

outcomes <- c("investor_return", "fair_1_repay", "max_1_repay")
outcome_labels <- c(
  investor_return = "Investor Cumulative Return",
  fair_1_repay = "Fair Agent Mean Repayment",
  max_1_repay = "Max Agent Mean Repayment"
)

ground_truth <- dgp_full %>%
  group_by(world) %>%
  summarise(across(all_of(outcomes), mean), .groups = "drop")

ate_true <- tibble(
  outcome = outcomes,
  label = outcome_labels[outcomes],
  ate_true = map_dbl(outcomes, function(y) {
    ground_truth[[y]][ground_truth$world == 1] -
      ground_truth[[y]][ground_truth$world == 0]
  })
)

kable(ate_true, digits = 4, col.names = c("Outcome", "Label", "True ATE"),
      caption = "Ground-truth ATEs from the DGP") %>%
  kable_styling(bootstrap_options = c("striped"), full_width = FALSE)
```

```{r gcomp-functions}
# ============================================================
# G-Computation Estimator
# ============================================================

gcomp_estimate <- function(data, outcome, covariates) {
  rhs <- paste(c("world", covariates), collapse = " + ")
  if (length(covariates) > 0) {
    interactions <- paste0("world:", covariates)
    rhs <- paste(c(rhs, interactions), collapse = " + ")
  }
  fml <- as.formula(paste(outcome, "~", rhs))
  fit <- lm(fml, data = data)

  d0 <- d1 <- data
  d0$world <- 0
  d1$world <- 1
  ate <- mean(predict(fit, newdata = d1) - predict(fit, newdata = d0))
  list(ate = ate, fit = fit)
}

# Define covariate sets
covs_A_candidates <- c("episode",
                        "eval_repay_fair", "eval_repay_max",
                        "eval_sd_fair", "eval_sd_max", "eval_invest",
                        paste0("fair_h", 1:5), paste0("max_h", 1:5))

# Keep only covariates that actually exist and aren't all-NA
covs_A <- covs_A_candidates[
  covs_A_candidates %in% names(dgp_full) &
    map_lgl(covs_A_candidates, ~ !all(is.na(dgp_full[[.x]])))
]
covs_B <- c("episode")

cat(sprintf("Setting A covariates (%d): %s\n",
            length(covs_A), paste(covs_A, collapse = ", ")))
cat(sprintf("Setting B covariates (%d): %s\n",
            length(covs_B), paste(covs_B, collapse = ", ")))
```

```{r point-estimates}
# ============================================================
# Point estimates (full sample)
# ============================================================

point_ests <- map_dfr(outcomes, function(y) {
  w1 <- mean(dgp_full[[y]][dgp_full$world == 1], na.rm = TRUE)
  w0 <- mean(dgp_full[[y]][dgp_full$world == 0], na.rm = TRUE)
  tibble(
    outcome = y,
    naive   = w1 - w0,
    gcomp_A = gcomp_estimate(dgp_full, y, covs_A)$ate,
    gcomp_B = gcomp_estimate(dgp_full, y, covs_B)$ate
  )
})

kable(point_ests, digits = 4,
      col.names = c("Outcome", "Naive DiM", "G-comp A (Full)", "G-comp B (Behav.)"),
      caption = "Point estimates from full sample") %>%
  kable_styling(bootstrap_options = c("striped"), full_width = FALSE)
```

```{r bootstrap, cache=TRUE}
# ============================================================
# Cluster Bootstrap (seed-level)
# ============================================================

all_seeds <- sort(unique(dgp_full$seed))
n_seeds <- length(all_seeds)

boot_one <- function(b) {
  boot_seeds <- sample(all_seeds, n_seeds, replace = TRUE)
  boot_data <- map2_dfr(boot_seeds, seq_along(boot_seeds), function(s, idx) {
    dgp_full %>% filter(seed == s) %>% mutate(boot_id = idx)
  })

  map_dfr(outcomes, function(y) {
    w1 <- mean(boot_data[[y]][boot_data$world == 1], na.rm = TRUE)
    w0 <- mean(boot_data[[y]][boot_data$world == 0], na.rm = TRUE)
    ate_naive <- w1 - w0

    ate_A <- tryCatch(
      gcomp_estimate(boot_data, y, covs_A)$ate,
      error = function(e) NA_real_
    )
    ate_B <- tryCatch(
      gcomp_estimate(boot_data, y, covs_B)$ate,
      error = function(e) NA_real_
    )

    tibble(boot = b, outcome = y,
           ate_naive = ate_naive, ate_A = ate_A, ate_B = ate_B)
  })
}

cat(sprintf("Running %d bootstrap replicates (seed-level clustering)...\n", N_BOOT))
t0 <- Sys.time()
boot_results <- map_dfr(seq_len(N_BOOT), function(b) {
  if (b %% 200 == 0) cat(sprintf("  [%d / %d]\n", b, N_BOOT))
  boot_one(b)
})
elapsed <- difftime(Sys.time(), t0, units = "mins")
cat(sprintf("Bootstrap complete in %.1f minutes.\n", as.numeric(elapsed)))
```

```{r eval-metrics}
# ============================================================
# Evaluation: Bias, RMSE, Coverage
# ============================================================

estimators <- c("ate_naive", "ate_A", "ate_B")
est_labels <- c(ate_naive = "Naive DiM",
                ate_A = "G-comp (Full Info)",
                ate_B = "G-comp (Behavioral Only)")

eval_estimator <- function(ates, truth, alpha = 0.05) {
  ates <- ates[!is.na(ates)]
  if (length(ates) == 0) return(tibble(
    bias = NA, rmse = NA, coverage = NA,
    mean_est = NA, se = NA, ci_lo = NA, ci_hi = NA))

  mn   <- mean(ates)
  se   <- sd(ates)
  bias <- mn - truth
  rmse <- sqrt(mean((ates - truth)^2))
  z    <- qnorm(1 - alpha / 2)

  # Coverage: fraction of bootstrap CIs (ate_b ± z*se) containing truth
  coverage <- mean((ates - z * se <= truth) & (ates + z * se >= truth))

  tibble(bias = bias, rmse = rmse, coverage = coverage,
         mean_est = mn, se = se,
         ci_lo = quantile(ates, alpha / 2),
         ci_hi = quantile(ates, 1 - alpha / 2))
}

eval_table <- map_dfr(outcomes, function(y) {
  truth <- ate_true$ate_true[ate_true$outcome == y]
  map_dfr(estimators, function(est) {
    ates <- boot_results %>% filter(outcome == y) %>% pull(!!sym(est))
    metrics <- eval_estimator(ates, truth, ALPHA)
    bind_cols(tibble(outcome = y, estimator = est, truth = truth), metrics)
  })
}) %>%
  mutate(
    outcome_label = outcome_labels[outcome],
    estimator_label = est_labels[estimator]
  )
```

## Evaluation Table

```{r eval-table-display}
eval_display <- eval_table %>%
  select(outcome_label, estimator_label, truth, mean_est,
         bias, rmse, coverage, se) %>%
  arrange(outcome_label, estimator_label)

kable(eval_display, digits = 4,
      col.names = c("Outcome", "Estimator", "True ATE", "Mean Est.",
                     "Bias", "RMSE", "Coverage", "SE"),
      caption = "Bias, RMSE, and 95\\% CI Coverage across estimators and outcomes") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE, font_size = 11) %>%
  column_spec(5, bold = TRUE)
```

## Bootstrap Distributions

```{r fig-bootstrap, fig.cap="Bootstrap distributions of ATE estimates. Red dashed line marks the ground-truth ATE.", fig.height=10}
boot_long <- boot_results %>%
  pivot_longer(starts_with("ate_"),
               names_to = "estimator", values_to = "ate_hat") %>%
  mutate(
    estimator_label = est_labels[estimator],
    outcome_label = outcome_labels[outcome]
  )

ggplot(boot_long, aes(x = ate_hat, fill = estimator_label)) +
  geom_histogram(alpha = 0.6, bins = 50, position = "identity") +
  geom_vline(
    data = ate_true %>% mutate(outcome_label = outcome_labels[outcome]),
    aes(xintercept = ate_true),
    color = "red", linetype = "dashed", linewidth = 1
  ) +
  facet_wrap(~ outcome_label, scales = "free", ncol = 1) +
  scale_fill_brewer(palette = "Set2") +
  labs(x = "Estimated ATE", y = "Count", fill = "Estimator") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom", strip.text = element_text(face = "bold"))
```

## Bias Comparison

```{r fig-bias, fig.cap="Bias by estimator and outcome. Dashed line at zero.", fig.height=8}
ggplot(eval_table, aes(x = estimator_label, y = bias, fill = estimator_label)) +
  geom_col(alpha = 0.8, width = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  facet_wrap(~ outcome_label, scales = "free_y", ncol = 1) +
  scale_fill_brewer(palette = "Set2") +
  labs(x = NULL, y = "Bias (mean estimate - true ATE)", fill = "Estimator") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 15, hjust = 1))
```

## RMSE Comparison

```{r fig-rmse, fig.cap="RMSE by estimator and outcome.", fig.height=8}
ggplot(eval_table, aes(x = estimator_label, y = rmse, fill = estimator_label)) +
  geom_col(alpha = 0.8, width = 0.6) +
  facet_wrap(~ outcome_label, scales = "free_y", ncol = 1) +
  scale_fill_brewer(palette = "Set2") +
  labs(x = NULL, y = "Root Mean Squared Error", fill = "Estimator") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 15, hjust = 1))
```

## Coverage

```{r fig-coverage, fig.cap="95\\% CI coverage. Red dashed line marks the nominal 95\\% level.", fig.height=8}
ggplot(eval_table, aes(x = estimator_label, y = coverage, fill = estimator_label)) +
  geom_col(alpha = 0.8, width = 0.6) +
  geom_hline(yintercept = 1 - ALPHA, linetype = "dashed", color = "red") +
  facet_wrap(~ outcome_label, ncol = 1) +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = NULL, y = "Coverage Rate", fill = "Estimator") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 15, hjust = 1))
```

## ATE by Training Episode

```{r fig-ate-episode, fig.cap="ATE by training episode, showing how the causal effect of communication evolves over the course of learning.", fig.height=10}
episode_means <- dgp_full %>%
  group_by(episode, world) %>%
  summarise(across(all_of(outcomes), mean), .groups = "drop")

ep_w0 <- episode_means %>% filter(world == 0)
ep_w1 <- episode_means %>% filter(world == 1)

episode_ates <- tibble(
  episode = ep_w0$episode,
  ate_investor = ep_w1$investor_return - ep_w0$investor_return,
  ate_fair = ep_w1$fair_1_repay - ep_w0$fair_1_repay,
  ate_max = ep_w1$max_1_repay - ep_w0$max_1_repay
) %>%
  pivot_longer(-episode, names_to = "outcome", values_to = "ate") %>%
  mutate(outcome_label = case_when(
    outcome == "ate_investor" ~ "Investor Cumulative Return",
    outcome == "ate_fair" ~ "Fair Agent Mean Repayment",
    outcome == "ate_max" ~ "Max Agent Mean Repayment"
  ))

ggplot(episode_ates, aes(x = episode, y = ate)) +
  geom_line(alpha = 0.3) +
  geom_smooth(method = "loess", se = TRUE, color = "steelblue", linewidth = 1.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  facet_wrap(~ outcome_label, scales = "free_y", ncol = 1) +
  labs(x = "Training Episode", y = "ATE (W1 - W0)") +
  theme_minimal(base_size = 12) +
  theme(strip.text = element_text(face = "bold"))
```

## Paired Seed Analysis

```{r fig-paired, fig.cap="Paired within-seed ATEs across all seeds. Each point is one seed; bar shows the paired difference.", fig.height=8}
seed_by_world <- dgp_full %>%
  group_by(seed, world) %>%
  summarise(across(all_of(outcomes), mean), .groups = "drop")

sw0 <- seed_by_world %>% filter(world == 0) %>% arrange(seed)
sw1 <- seed_by_world %>% filter(world == 1) %>% arrange(seed)

seed_means <- tibble(
  seed = sw0$seed,
  diff_investor = sw1$investor_return - sw0$investor_return,
  diff_fair = sw1$fair_1_repay - sw0$fair_1_repay,
  diff_max = sw1$max_1_repay - sw0$max_1_repay
)

seed_long <- seed_means %>%
  pivot_longer(-seed, names_to = "outcome", values_to = "diff") %>%
  mutate(outcome_label = case_when(
    outcome == "diff_investor" ~ "Investor Cumulative Return",
    outcome == "diff_fair" ~ "Fair Agent Mean Repayment",
    outcome == "diff_max" ~ "Max Agent Mean Repayment"
  ))

ggplot(seed_long, aes(x = factor(seed), y = diff)) +
  geom_col(fill = "steelblue", alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  facet_wrap(~ outcome_label, scales = "free_y", ncol = 1) +
  labs(x = "Seed", y = "Within-Seed Difference (W1 - W0)") +
  theme_minimal(base_size = 11) +
  theme(strip.text = element_text(face = "bold"),
        axis.text.x = element_text(size = 7))
```


# Discussion

```{r discussion-stats, include=FALSE}
# Prepare some numbers for inline reporting
naive_inv_bias <- eval_table %>%
  filter(outcome == "investor_return", estimator == "ate_naive") %>%
  pull(bias)
gcomp_a_inv_bias <- eval_table %>%
  filter(outcome == "investor_return", estimator == "ate_A") %>%
  pull(bias)
gcomp_b_inv_bias <- eval_table %>%
  filter(outcome == "investor_return", estimator == "ate_B") %>%
  pull(bias)

n_seeds_pos <- sum(seed_means$diff_investor > 0)
n_seeds_neg <- sum(seed_means$diff_investor < 0)
```

The simulation results reveal several patterns relevant to the choice of
causal inference methodology in multi-agent settings.

**Effect size.** The ground-truth ATE of inter-agent communication on investor
returns is modest (`r sprintf("%.1f", ate_true$ate_true[1])`), consistent with
the hypothesis that in the two-agent case, the primary interference channel
operates indirectly through the shared investor rather than through direct
agent-to-agent coordination. The paired seed analysis shows substantial
heterogeneity: `r n_seeds_pos` seeds show positive effects and `r n_seeds_neg`
show negative effects, suggesting that the direction of the communication effect
depends on the specific learning trajectory realized under each seed.

**Estimator performance.** [Interpret based on actual results after knitting.]

**Setting A vs. Setting B.** [Compare bias, RMSE, and coverage between the
full-information and behavioral-only g-computation settings.]

**Non-stationarity.** The ATE-by-episode figure illustrates how the causal
effect evolves over the course of learning. Early episodes, where agents are
still exploring ($\epsilon$ is high), may show different treatment effects than
late episodes where policies have converged. This non-stationarity is a genuine
feature of the DGP --- not an artifact --- and highlights why sequential methods
like g-computation are appropriate: they can condition on episode number as a
proxy for learning stage.

**Limitations.** The Setting A covariates are run-level summaries extracted from
the evaluation simulation, not true per-episode hidden states from the training
loop. This means Setting A captures *between-run* variation in what the RNN
learned, but not *within-run* episode-to-episode hidden state dynamics. A fully
realized Setting A would require logging $h_t$ at each training episode, which
is a straightforward engineering extension but was not implemented for this
project. Consequently, the Setting A vs. Setting B contrast may understate the
true benefit of observing hidden states.

**Implications.** In adversarial multi-agent environments, the choice of causal
estimator should be guided by what the analyst can observe. When internal agent
states are accessible (as in simulation), g-computation with full conditioning
provides unbiased estimates. When only behavioral traces are available (as in
real human experiments), the analyst faces a bias--variance tradeoff that depends
on the strength of the latent confounding.


# References

1. Dezfouli, A., Nock, R., & Dayan, P. (2020). Adversarial vulnerabilities
   of human decision-making. *PNAS*, 117(46), 29221--29228.

2. Robins, J. M. (1986). A new approach to causal inference in mortality
   studies with a sustained exposure period. *Mathematical Modelling*,
   7(9-12), 1393--1512.

3. Bongers, S., Forré, P., Peters, J., & Mooij, J. M. (2021). Foundations
   of structural causal models with cycles and latent variables. *Annals of
   Statistics*, 49(5), 2885--2915.

4. Nobandegani, A. S., Rish, I., & Shultz, T. R. (2023). Towards machines
   that trust: AI agents learn to trust in the trust game. *arXiv:2312.12868*.

5. Park, S., Lu, W., & Yang, S. (2025). Evaluating and learning optimal
   dynamic treatment regimes under truncation by death. *NeurIPS 2025*.
